---
title: "Cacao Data Tidy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Environment
```{r environ, echo=TRUE, message=FALSE, warning=FALSE}

library(timeSeries)
library(reshape2)
library(cluster)
library(gplots)
library(ggfortify)
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
library(base)
library(graphics)
library(stats)
library(fBasics)
library(pvclust)
library(pca3d)
library(randomForest)
library(varSelRF)
library(e1071)
library(VennDiagram)
library(googlesheets4)
```

# 2.Data description and acquisition 

This data comes from the LC/MS TOF based untargeted metabolomics of cacao seed colected from three different regions of Colombia: Bajo calima (Bcal), Cisneros (Cis), Sabaletas (Sab). From each region the metabolomic profile was performed under three different fermentation time (72, 96, and 120 h)   

```{r acquisition}
raw.data <- read_csv(file = "Data/Buenaventura05222020.csv")
head(raw.data)
```

# 3. Data tidy
For data tidying we have to take into the account main parameters:
1. Check character and negative values and replace it by cero
2. Peak width: all masses that have a peak width less than 3 sec should be filtered.
3. The masses must be present in the 90% of all the samples
4. The intensity of the masses: It will depend of the noise threshold. In some cases was 500 units of intensity
5. The remaining NAs will be replaced by the baseline signal (minimum value != 0)

Taking in to the account that the data is already filtered and normalized I will omit step 2 and 4

## 3.1. Data tidy Step 2. 
```{r tidy2}
# Extract numeric df
df <- raw.data %>% 
  select(4:4411) 

# Convert caracter values to NA  
df.1 <- 
  df %>%  
  map_df(as.character) %>% 
  map_df(as.numeric)

# identifying negative data in the df
df.1 %>% 
  summarize(pos = sum(.>0, na.rm = T),
            neg = sum(.<0, na.rm = T), 
            zero = sum(. == 0, na.rm = T),
            na = sum(is.na(.)))

# Convert NA to Zero
df.1[is.na(df.1)] <- 0


# Step 2. I will remove the masses that that have zero value in the 90% of the samples
data.WO_zero <- df.1[, colSums(df.1 == 0)/nrow(df.1) < .9, drop = FALSE]

```

## 3.2. Data tidy Step 5. 
The missing values and zero will be replaced with the minimum value. Assuming the minimum as the baseline. 

```{r tidy4}
# Step 5.
data.WO_zero[data.WO_zero <= 0] <- NA
minimum <- min(data.WO_zero, na.rm = T)
```

```{r}
min.fill <- data.WO_zero
min.fill[is.na(min.fill)] <- minimum

min.fill %>% 
  summarize(pos = sum(.>0, na.rm = T),
            neg = sum(.<0, na.rm = T), 
            zero = sum(. == 0, na.rm = T),
            na = sum(is.na(.)),
            min = min(.))

write.csv(min.fill, file = "Data/cleanData.csv")
```
